\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-08-05 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement texte modèle encore

Conditional on exporting abroad, the profit-maximizing program on the foreign market requires the inclusion of variable international trade costs according to:
\begin{eqnarray*}
\max_{p^x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
\text{s.t. }&&c^x(\varphi) = \left(\frac{p^x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
\end{eqnarray*}

\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p^x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market. Solving for this yields:
$$p^x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right]$$

@@

To understand the changes induced by the additive costs, let us start from reasoning in the benchmark model with only ad-valorem costs of \cite{melitz}. In this case, the reduction in transport costs induces an increase in selection on the domestic market ($\phi^\ast$ increases, Figure \textbf{XXXX STOP HERE }), but more domestic firms can now export ($\phi_x^\ast$ decreases). Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases). As a whole, the total number of varieties accessible to the consumer reduces ($M_T$ decreases slightly). Everything else equal, this reduces welfare. This is counteracted by a price competitiveness effect. Due to the tougher selection on the domestic market, the less productive firms no longer enter. As a result, the average productivity on the domestic market increases ($\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase). This price competitiveness effect pushes welfare upwards;

Reduction in additive costs: More negative variety effect

>>>>

Conditional on exporting abroad, the profit-maximizing program on the foreign market requires the inclusion of variable international trade costs according to:
\begin{eqnarray*}
\max_{p_x(\varphi)} \pi_x(\varphi) &=& \left(p^x(\varphi) -\frac{\tau}{\varphi}  \right)c^x(\varphi) - tc^x(\varphi) \\
\text{s.t. }&&c^x(\varphi) = \left(\frac{p_x(\varphi)}{P^\ast}  \right)^{-\sigma} \frac{E^\ast}{P^\ast}
\end{eqnarray*}

\noindent where $E^\ast$ and $P^\ast$ refer to nominal expenditures and the price index in the foreign country, $c^x(\omega)$ to the demand for variety $\omega$ from the Foreign household and $p_x(\omega)$ to the price set by the firm producing the variety $\omega$ on the export market. Solving for this yields:
\begin{equation}
p_x(\varphi) = \frac{\sigma}{\sigma-1}\left[\frac{\tau}{\varphi} +t \right] \label{eq:px}
\end{equation}

@@

To understand the changes induced by the additive costs, consider first the benchmark model with only ad-valorem costs, as in \cite{melitz}. In this case, the reduction in transport costs induces an increase in selection on the domestic market, but more domestic firms can now export (Figure \ref{fig:competitivity_air_decomp} for Air transport). Accordingly, the number of domestic active firms reduces, even if a larger number of them is able to export ($M$ decreases but $M_x$ increases, see Figure \ref{fig:variety_air_decomp}). As a whole, the total number of varieties accessible to the consumer reduces (Figure \ref{fig:variety_air_decomp}, $M_T$ decreases slightly). Everything else equal, this reduces welfare (Figure \ref{fig:welfare_decomp}). Yet, this detrimental effect is counteracted by a price competitiveness effect. Due to the tougher selection on the domestic market, the less productive firms no longer enter. As a result, the average productivity on the domestic market increases (Figure \ref{fig:competitivity_air_decomp}, $\widetilde{\varphi}^\ast$ increases), which pushes the average productivity upwards ($\widetilde{\varphi}^\ast$ and $\widetilde{\varphi}_T$ increase). As reported in Figure \ref{fig:welfare_decomp}, the strength of the price competitiveness effect dominates, inducing a increase in welfare following the reduction in iceberg costs, consistently with the analytical results of \cite{melitz}.

What does it change when the reduction in international transport costs also occurs through a reduction in its additive component? As reported in Figure \ref{fig:welfare_decomp} (right grey bars), trade liberalisation still induces welfare gains, and even larger than in the pure ad-valorem case. This is consistent with the analytical results of \cite{sorensen2014}, in a different exercise of trade liberalization though. \footnote{\cite{sorensen2014} compares the welfare effect of a trade costs reduction depending whether it is exclusively ad-valorem or exclusively additive, the comparison between both trade cost decreases being a criterion of equal impact on trade openness. Our quantitative model offers a more general and flexible set-up, as we can virtually consider all cases of trade liberalization. Based on our empirical estimates, we compare the effects of the same decrease in overall transport costs depending on 1) whether additive costs are present or not, and 2) when they are present, assuming that the trade cost decrease is achieved by a reduction of both additive and ad-valorem components.} Quantitatively, the effects are sizeable.

Through which mechanism does it occur? As underlined by \cite{sorensen2014}, a reduction in the additive cost reduces the export-market price of the high-productivity exporters relative to the low-productive ones (this can be shown from Equation (\ref{eq:px})). Within the exporters category, this favors the more productive firms, whose market share, e.g. revenues increases relatively more, as can be shown from Equation (\ref{eq:rx}). Everything else equal, this pushes the average productivity on the export market upwards. In contrast to ad-valorem costs, the reduction in additive costs induces a \textit{within-exporters selection effect}, in favor of the more productive firms. At the general equilibrium, this within-exporters selection effect also affects the \textit{classic} selection effect at the extensive margin, at the heart of the \cite{melitz} model: As the expected export profit increases, it is necessary to be more productive to satisfy the free-entry condition: Everything else equal,
